---
title: "Numerical Integration"
subtitle: "Numerical Introductory Course"
author: "Marvin Gauer (580553)"
header-includes:
- \usepackage{titling}
- \pretitle{\begin{center}\LARGE\includegraphics[width=5cm]{husiegel_bw.png}\\[\bigskipamount]}
- \title{\vspace{3cm} \Huge Numerical Integration \vspace{3cm}}
- \preauthor{\centering \Large Numerical Introductory Course \\}
- \author{\Large Marvin Gauer (580553)}
- \posttitle{\end{center}}

output: pdf_document
citation_package: natbib
bibliography: biblio.bib
---
```{r include = TRUE, echo = FALSE, comment = NA}
knitr::opts_chunk$set(fig.width=3, fig.height=3)
```
```{r include = FALSE}
rm(list=ls())

library(ggplot2)

```
\thispagestyle{empty}
\clearpage
\thispagestyle{empty}
\tableofcontents
\newpage
\pagenumbering{arabic}
\setcounter{page}{1}

# 1. Abstract

# 1. Introduction

# 1. Motivation

Integration is an important operation in mathematics. Unfortunately, in real life applications one might find it extremely difficult or even impossible to solve certain integrals. Due to the continous improvement in computational power one might address this issue by numercially approximating the integral of interest. In order to do so, several procedures have been developed, each with it's own pros and cons. One may for example use the "Rechteckappromation", the "Trapezapproximation"
or the in this paper elaborated on Monte Carlo Method. 

# 2. Literature Review

lalalala

# 3. Theory

In the following section I want to explain some of the most popular methods in numerical integration. These can be distingushed into one and multi-dimensional methods. Furthermore one might distinguesh numerical integration methods further into deterministic and probabilistic methods.

# 3.1 Review: The Riemann Integral



# 3.2 One-Dimensional Procedures

lalalala

# 3.3 Multi-Dimensional Procedures

lalalala

# 3.4 Integrals over infinite Intervals

lalalala

# 4. Application: Approximation of the Normal Distribution

lalalala

# 5. Conclusion

lalalala

# 6. Bibliography

@Clausthal





